---
title: "Summary Playground"
author: "Kimberly Bastille, Modified by WK"
date: "5/11/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(here)
library(dplyr)

shade.alpha <- 0.3
shade.fill <- "lightgrey"
lwd <- 1
pcex <- 2
trend.alpha <- 0.5
trend.size <- 2
hline.size <- 1
hline.alpha <- 0.35
hline.lty <- "dashed"
label.size <- 5
hjust.label <- 1.5
letter_size <- 4
ylim_re <- c(5e6, 30e6)
ylim_rd <- c(1.75,2.75)
ylim_ra  <- c(0, 2e6)


```


```{r data-list}
datalist<-c("dataA", "dataB", "dataC", "dataD", "dataE","dataF")
#datalist<-c("billion", "Chla", "SST")

```


```{r functions-to-write-into-package}

#need to subset all commands by region
summarize_me_for_fun<- function(data){
  dat<-read.csv(here::here("data-raw", paste0(data, ".csv")))
  # Summary Stats
  # TRange <- quantile(dat$y, c(.025,.05, .25, .75,.95,.975))
  # TMarker <- round(mean(dat$y),digits=5)
  # TName <- dat$title
  # TMin <- min(dat$y)
  # TRange <- cbind(t(TRange),TMarker, TMin, TName)
  # colnames(TRange)<-c("Range1","Range2","Range3","Range4",
  #                      "Range5","Range6","Marker",
  #                      "Value","Min")
  # 
  ## Status
  mn<- mean(dat$y, na.rm=TRUE)
  last5<- dat %>% dplyr::filter(x %in% c(max(dat$x):(max(dat$x) - 4))) 
  mn5 = mean(last5$y, na.rm=TRUE)            # mean over the eval period
  status <- ifelse(mn5>mn, paste0("greater than"), paste0("less than"))
  
  ##Trend
  m1 = lm(last5$x~last5$y)
  s1 <- summary(m1)
  b1 <- s1$coefficients[2,1]
  pval <- s1$coefficients[2,4]
  pred = predict(m1)
  delta = pred[length(pred)] - pred[1]  # gives magnitude and direction of change over eval period
  Z = abs(delta)-(sd(dat$y, na.rm=T))
  trend <- ifelse(Z > 0, paste0("significantly"), paste0(""))
  ## Guage
  cdf <- stats::ecdf(dat$y)
  guage <- round(cdf(mn5) *100)
  
  summary_df<- data.frame(title = unique(dat$title),
                          region = unique(dat$region),
                          #TRange, 
                          mn, 
                          mn5,
                          status, 
                          trend,
                          delta,
                          Z,
                          guage)
  return(summary_df)
}


#way to loop ggplot to create individual plots per region?
plot_me_for_fun<- function(data){
  dat<-read.csv(here::here("data-raw", paste0(data, ".csv")))
  pt<-dat %>% ggplot2::ggplot(aes(x = x, y = y))+
    ggplot2::annotate("rect", fill = "green", alpha = 0.3,
                      xmin = max(dat$x - 10), xmax = max(dat$x),
                      ymin = -Inf, ymax = Inf)+
    theme(panel.background = element_rect(fill = "transparent"), # bg of the panel
          panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
    ggplot2::geom_line(color = "red")+
    ggplot2::geom_point(color = "red")+
    ggplot2::geom_hline(aes(yintercept = mean(y),),
               size = hline.size,
               alpha = hline.alpha,
               linetype = hline.lty) +
    #geom_hline(aes(yintercept = sd_TEMP,),
    ggplot2::geom_hline(aes(yintercept = mean(y) + sd(y),),
               size = hline.size,
               alpha = hline.alpha,
               linetype = "dotted") +
    #geom_hline(aes(yintercept = -sd_TEMP,),
    ggplot2::geom_hline(aes(yintercept = mean(y)-sd(y),),
               size = hline.size,
               alpha = hline.alpha,
               linetype = "dotted") +
    ggplot2::ggtitle(paste(dat$title),subtitle=(paste(dat$region)))+
    ggplot2::ylab(expression("y"))+ ## Fix this with Will
    ggplot2::xlab(expression("Year"))
  print(pt)
}
```

## Plots

These are all the plots you build - look how pretty. 

```{r plots, warning=FALSE}
for (i in datalist) {
  plot_me_for_fun(i)
}
```

## Summary data

```{r summary, warning=FALSE}
summarydf<- data.frame() #blank df to be populated

for (i in datalist) {
  dat<-summarize_me_for_fun(i)
  summarydf<-summarydf %>% rbind(dat)
}

DT::datatable(summarydf)

write.csv(summarydf, file = here::here("data","summarydf.csv"))
```


### Will's spitballs 5/11/23
- Need to find way to batch read in and reformat existing EIWG data
```{r}
# library(tidyr)
# library(here)
# library(dplyr)
# 
# #test data
# #load in data, skip 2 rows (y label, title)
# SST <- read_csv("data-eiwg/SST.csv", skip = 2)
# 
# SST_long <- gather(SST, region, degC, 2:10, factor_key=TRUE)
# SST_format <- SST_long %>% mutate(title = "SST")
# colnames(SST_format) <- c('y','region','x','title')
# head(SST_format)
# 
# #test batch
# #take a swing at loop
# 
# eiwg_files <- list.files(here::here("data-eiwg"))
# for(i in eiwg_files) {                        
#   name <- gsub(".csv","",i)
#   assign(name,                                   
#   read.csv(here("data-eiwg",i), skip = 2))
# }
# #this^ works but you lose units
# 
# indicator_names <- gsub(".csv","",eiwg_files)
# dfs <- Filter(function(x) is(x, "data.frame"), mget(ls()))
# res<- lapply(dfs, function(w) {gather(w, region, unit, 2:ncol(w), factor_key=TRUE)})
# indicators <- mapply(cbind, res, "title"=indicator_names, SIMPLIFY=F)
# 
# #produces list of indicator dfs in same format as dummy data
```

### Notes for improvement 3/29/23

- Guage read out - Percentile rank of last 5 compared to mean of whole series. 

```{r}
# mn5 = mean(Y5, na.rm=TRUE)            # mean over the eval period
#   mn = mean(co)                         # mean over entire data series
#  # pTileRankNew <- round(((length(co_all[co_all <= mn5])/(length(co_all)+1))*100)) # older version of percentile
#   cdf <- ecdf(co_all)
#  pTileRankECDF <- round(cdf(mn5) *100)
```

- Trend - Is mean of last 5 years above or below 1 SD?

```{r}
# if (trendAnalysis==T) {
#     par(mar=c(2.5,0,3,0))
#     plot(1, xlim=c(0.94,1.06), ylim=c(0.6, 1.6), col=0, axes=F, xlab="", ylab="")
# 
#     # Mean of eval period outside 10th or 90th percentile?
#     points(1, 1.225, pch=20, cex=5)
#     maxNA <- length(Y5) * 0.4 # scale the maximum allowable NAs in eval period
#     print(paste("Max Na's for eval period= ",maxNA))
#     if (sum(is.na(Y5)) < maxNA)  {
#       if (mn5 >= ptile[3])  { text(1, 1.225, col="white", "+", cex=2.6, font=2) }
#       if (mn5 >= ptile[3])  { text(1, 1.225, col="white", "+", cex=2.6, font=2) }
#       if (mn5 <= ptile[1])  { text(1, 1.225, col="white", "-", cex=2.6, font=2) }
# }}
#       # Change over last 5 yrs > 1 s.d.?
#       m1 = lm(Y5~X5)
#       s1 <- summary(m1)
#       b1 <- s1$coefficients[2,1]
#       pval <- s1$coefficients[2,4]
#       pred = predict(m1)
#       delta = pred[length(pred)] - pred[1]  # gives magnitude and direction of change over eval period
#       Z = abs(delta)-(sd(co_all, na.rm=T))
# Z positive means the total change over eval period exceeds 1 s.d. of entire series
```

Z if pos use significantly
delta says if it above or below mean


LME column 
Indicator title column
Y-label column
